Observation of the orientation of snow hydrometeors at sheltered and unsheltered sites

### 2024-12-11 Review_R1_answers_202412.tex

\noindent

\begin{table}[H]
\centering
\begin{tabular}{ |c|c| } 
 \hline
 {\itshape italic text  } & {\itshape Comments of the reviewers}  \\ 
 roman text               & Our answers or comments   \\ 
 \blue{blue text}         & \blue{Verbatim quotes from the manuscript}  \\ 
 \hline
\end{tabular}
\end{table}

\noindent
A manuscript is included with highlighted differences and modifications applied during revision is submitted as well.

\subsection*{Editor Comment}
\textit{Thank you for submitting "Observation of falling snowflakes orientation at sheltered and unsheltered sites" [Paper \#2024GL113042] to Geophysical Research Letters. I have received 2 reviews of your manuscript, which are included below and/or attached. As you can see, the reviews indicate that major revisions are needed before we can consider proceeding with your paper. I am therefore returning the paper to you so that you can make the necessary changes.
}\\

\noindent
In this document we address point-by-point the remarks of the two reviewers and make the necessary changes in the manuscript.

@@

\newpage
\subsection*{Reviewer 1}

\textit{The manuscript is well-written. The authors did a great job addressing the observational issues with the Multi-Angel Snowflake Camera (MASC). I particularly appreciate the reverse-engineering numerical experiments: idealized and realistic objects were generated to evaluate various methods for describing the orientation of those objects. The findings from the numerical experiment, together with the analysis of the in-situ data, will be a valuable addition to the literature and a useful resource for the remote sensing community. I recommend the manuscript be published in GRL, after the following issues are addressed:}\\

\noindent
We thank the Reviewer for their positive feedback and for the suggestions and comments listed below.

@@

\subsubsection*{Major comments:}
\begin{enumerate}
    \item \textit{The discussion about the impact of the environmental winds on the orientation of the snowflakes is not sound. Wind speed alone may not be sufficient to capture this impact. The distribution of wind components likely plays a more significant role.}

    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have co-located wind speed information available).

    We carefully revised the manuscript to make sure not to overstate the role of wind speed alone and also to clarify when we cannot separate between the "natural" contribution of wind on the distribution of orientations and the "artificial" contribution due to the turbulence generated by strong wind interacting with the hollow measurement system of the MASC. A few changes are highlighted in the manuscript. We may cite for example in Sec.~3:

@@

\subsubsection*{Comments:}

\begin{enumerate}

    \item \textit{There is a very recent laboratory study about ice particle orientation which should be added to reference list and I also highly suggest to discuss their results. They appear to be slightly contradictory to your study:} \cite{Stout_ACP_2024} \\

    We thank the reviewer for this relevant, recommendation.  This high quality study, with laboratory experiments using snowflake replicas in a liquid solution shows that planar crystals may be falling with regimes that do not have a preferential orientation around 0$^\circ$. We believe that, given some of the assumptions of~\cite{Stout_ACP_2024}, their results are somehow complementary to our work and point out to the need to better understand the role of turbulence in orientation of snow particles. Regarding the cited study, we have the following comments:
    \begin{itemize}
        \item A relevant limitation of~\cite{Stout_ACP_2024} work is that, quoting, "it  only considers quiescent conditions, as the authors want to know under what conditions particle instability still occurs, even without the addition of turbulence.". This is quite different with respect to our field-based experiments.
        \item The study focuses on planar crystals and it uses 12 sub-types of this  hydrometeors (from plates to dendrites). In our case, the planar crystals are all classified together into an individual class, which is also one of the most difficult to classify with the MASC and to observe (given the 35 microns resolution of the MASC itself). In this sense our study is coarser and cannot go to this level of detail, which could be a reason why we do not observe such behavior. 
        \item Interestingly, another recent study~\cite{Koebschall_EF_2023} using replicas of snow aggregates and a liquid solution, showed instead that the particles tend to orient such that the area projected in the direction of flow is maximized, which is the type of behavior that is consistent with our observations. This is very interesting as two separate indoor studies, with similar methods, seem to indicate that individual crystals of planar type and aggregate snowflakes may be behaving differently regarding preferential orientation.
        \item It is unfortunate that the study reports, as possible backup evidence and as driving motivation, that previous studies using MASC also showed off-zero preferential orientations. This is exactly what we showed, in our paper, to be due to the bias of averaging \textbf{absolute value} of orientation estimates taken from different viewpoints and it was one of the driving motivation of our publication.  
        \item A further limitation is that the study considers objects falling in a liquid, which implies a far lower density ratio compared to hydrometeors in air, which in turn can impact the settling dynamics as discussed in recent studies by one of our coauthors (see~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}).
    \end{itemize}

    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}. At first, we include this study in Table~1. Then, in the Introduction, we mention this study as:\\

    \blue{ A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}. It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation. This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.  }\\

@@

    \noindent
    We agree with the reviewer, and we implemented their suggestion. \\

    \item \textit{Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak. Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more. It is certainly true that most radar polarimetric quantities are sensitive to orientation (L. 50-52). But this is by far not all in my opinion. Particle orientation also impacts the scattering signal in passive microwave (MW) observations (both from space and from ground). Many space-borne MW sensors are also polarization-sensitive. Considering the enormous efforts undertaken for example by ECMWF and other services to assimilate those all-sky MW observations, this aspect should be mentioned and cited. Another aspect is that an ice particle falling with a preferred orientation will have a very different sedimentation velocity. So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model). Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.}

    We thank the reviewer for the suggestion. While we left the second part of the Introduction a bit more "specialized" (targeting an audience interested in snow imagers and their data processing), we realized that we did not provide a satasfactory broad overview, neglecting all the implications listed by the reviewer. We rephrased therefore the first part of the Introduction, removed a few technical paragraphs and included most of the suggestions of the reviewer. Some new paragraphs now read:

    \blue{The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}. }

@@

    \blue{The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.}\\

    \item \textit{L. 46: It would be good to add a reference for this statement. I question a bit whether this general statement is actually true. I completely agree for plate-like crystals or large aggregates. But what is for example about bullet rosettes in high cirrus clouds? Or columns/needles? Do we have solid knowledge about their falling orientation? Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}

    We believe the reviewer is right. We were biased tp think mostly in terms of larger snow particles (snow aggregates, graupel) or planar crystals, so our formulation was overstating the current knowledge. We rephrased fully this sentence as:\\

    \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.  Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes. }\\

@@

    \item \textit{L. 47: What do you mean by "relatively simple"? Personally, I find all those ice particles terribly complicated.}\\

    We agree that this sentence needed to be rephrased (see previous point).

### 2024-12-11 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

\section{Introduction}
Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.

The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
%Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
The concept of orientation is linked to the anisotropic geometry of most snow particles and loses any meaning for isotropic, spherical particles. In the case of snowfall, the hydrometeors are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence). They report as well non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles. A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}. It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation. This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.

The problem of observing the orientations of hydrometeors while they fall in natural outdoor conditions, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

@@

We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}. The database includes observations collected in various sites worldwide, for example in the Swiss Alps, in Korea and in Antarctica between 2014 and 2024. The database includes both observations collected in conditions of natural wind exposure and, for some campaigns, with the instrument installed within a Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence and mean wind .

@@

MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as example) and pre-computed geometrical and microphysical descriptors. For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport. All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results. The resulting filtered dataset comprises 412'150 particles.

@@

Given that the measurements within DFIRs are less perturbed, we focus now on this siting type. With the original dataset's large size, we can conduct further data stratification and still keep a large data sample. Figure~\ref{fig:hydroclass} illustrates orientation distributions under two contrasting wind regimes ($<4$ ms$^{-1}$ and $>8$ ms$^{-1}$), stratified by hydrometeor type~\cite{Praz_AMT_2017} and average axis ratio. Stronger winds lead to broader and flatter distributions, where a clear mode around 0$^\circ$ orientation values is less evident. For elongated columnar and planar crystals (low axis ratio), this results in distributions showing off-zero modal values, which were not observed under low-wind conditions. A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles~\cite{Saffman_JFM_1956}. The results, especially for planar crystal types, must be considered with extreme care. In fact, the classification method of~\cite{Praz_AMT_2017}, which is trained on a supervised set, is able to identify planar crystals when their basal facet is facing towards the MASC cameras, making them recognizable.  This introduces a bias as some orientations would not allow for classification. For example, a thin plate falling perfectly horizontal, would appear in MASC images as a thin column.

@@

\section{Conclusions}
This study focused on the estimation of the orientation of falling snowflakes using a MASC, providing observational back-up to commonly made assumptions on orientation distributions. It showed how some common data processing choices (for the data processing of MASC data) can lead to biased estimates, and highlighted the effect of unsheltered sites on the collected measurements.
The main results are summarized as follows. 
At first, estimating orientation as the mean of the orientation of the best-fit ellipsoid from each camera view represents a reasonable approach in terms of minimizing errors and accurately reproducing the distribution of (3D) orientation values. Single-camera estimates exhibit high uncertainty compared to combining estimates from all three views and this should be kept in mind if single-view imagers are used. Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies (in natural outdoor conditions) are likely erroneous because of the artificially non-zero mode resulting from averaging the \textit{absolute value} of orientation estimates from each camera view. This explanation, supported by numerical simulations in the present study, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations. It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiralling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions. With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals.

>>>>

\section{Introduction}
Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.

The orientation and falling regime also affects the settling velocity of ice phase hydrometeors, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
%Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
The concept of orientation is linked to the anisotropic geometry of most snow particles and loses any meaning for isotropic, spherical particles. In the case of snowfall, the hydrometeors are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented). Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that there are differences due to the type of hydrometeor and environmental conditions (wind and turbulence). They also report non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles. A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}. It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation. This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.

The problem of observing the orientations of hydrometeors while they fall in natural outdoor conditions, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote sensing instruments are also commonly used, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

@@

We used a large data set of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}. The database includes observations collected in various sites worldwide, for example in the Swiss Alps, in Korea and in Antarctica between 2014 and 2024. The database includes both observations collected in conditions of natural wind exposure and, for some campaigns, with the instrument installed within a Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence and mean wind .

@@

MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as an example) and precomputed geometrical and microphysical descriptors. For the purpose of the study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport. All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results. The resulting filtered dataset comprises 412'150 particles.

@@

Given that the measurements within DFIRs are less perturbed, we focus now on this siting type. With the original dataset's large size, we can conduct further data stratification and still keep a large data sample. Figure~\ref{fig:hydroclass} illustrates orientation distributions under two contrasting wind regimes ($<4$ ms$^{-1}$ and $>8$ ms$^{-1}$), stratified by hydrometeor type~\cite{Praz_AMT_2017} and average axis ratio. Stronger winds lead to broader and flatter distributions, where a clear mode around 0$^\circ$ orientation values is less evident. For elongated columnar and planar crystals (low axis ratio), this results in distributions showing off-zero modal values, which were not observed under low-wind conditions. A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles. The results, especially for planar crystal types, must be considered with extreme care. In fact, the classification method of~\cite{Praz_AMT_2017}, which is trained on a supervised set, is able to identify planar crystals when their basal facet is facing towards the MASC cameras, making them recognizable.  This introduces a bias as some orientations would not allow for classification. For example, a thin plate falling perfectly horizontal, would appear in MASC images as a thin column.

@@

\section{Conclusions}
This study focused on the estimation of the orientation of falling snowflakes using a MASC, providing observational back-up to commonly made assumptions on orientation distributions. It showed how some common data processing choices (for the data processing of MASC data) can lead to biased estimates, and highlighted the effect of unsheltered sites on the collected measurements.
The main results are summarized as follows. 
At first, estimating orientation as the mean of the orientation of the best-fit ellipsoid from each camera view represents a reasonable approach in terms of minimizing errors and accurately reproducing the distribution of (3D) orientation values. Single-camera estimates exhibit high uncertainty compared to combining estimates from all three views and this should be kept in mind if single-view imagers are used. Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies (in natural outdoor conditions) are likely erroneous because of the artificially non-zero mode resulting from averaging the \textit{absolute value} of orientation estimates from each camera view. This explanation, supported by numerical simulations in the present study, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations. It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiraling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions. With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals.
